School accessibility: a new index applied to the case of São Paulo

### 2015-09-29 paper.Rmd
Final version abstract JoTG

---
title: 'School accessibility: a new index applied to enrollment and performance inequalities
  in Sao Paulo'
author: "Ana Moreno and Robin Lovelace"
output: pdf_document
bibliography: school-access-bib.bib
---

@@

Travel to school is an everyday reality for millions of young people around the world.
The mode, duration, safety, comfort and pollution levels of this trip has huge
impacts on the future generation, yet has received relatively little academic
attention.
Access to high quality quality schools is highly variable across geographical
space, making travel options are vital for ensuring a more equitable supply of
educational opportunity to diverse groups. Conversely, poor accessibility
to deprived area can reinforce social inequalities, with long-term implications.
Based on this emerging context, this paper develops a school accessibility index
for local areas.
This index is based on current travel patterns, local travel infrastructure serving the school
catchment and the provision of travel services such as public transport links.
Using data from the Sao Paulo Household Travel Survey (2007 and 2012), School Census
and electronic card data, we demonstrate how this index can be used
measure current provision and evaluate future interventions.
The results suggest that that school accessibility can
impact the social background of school attendees and
academic performance.
We conclude that extending the concept of local accessibility indicators
to the 'school commute' can ensure that transport decision makers take
equality and health impacts on young people into account.

# Introduction

@@

@@

This early work made the distinction between accessibility indeces that apply to zones or single 'desire lines':
"relative accessibility is defined as a measure of the effort of overcoming spatial separation between two points, while the integral accessibility is defined as a measure of the effort of overcoming spatial separation between a point and all other points within an area"
[@allen1993accesibility].

@@

>>>>

---
title: 'School accessibility: a new index applied to the case of São Paulo'
author: Ana I. Moreno-Monroy (Universitat Rovira i Virgili), Frederico R. Ramos (Fundação
  Getulio Vargas) and Robin Lovelace (University of Leeds)
output: word_document
bibliography: school-access-bib.bib
---

@@

A skewed distribution of schools towards wealthy areas can result in a mismatch between the provision of educational opportunity and the need for access to education.
This mismatch can be further reinforced by poor transport provision in lower income areas.
In this paper, we develop a school accessibility index that captures both the unequal distribution of schools and transport services across space.
Our index combines information on the spatial distribution of children and adolescents by income categories, the location of existing schools, the travel patterns of students, and the travel infrastructure serving the school catchment area into a single measure of school accesibility. We use our index to measure school accessibility by areas across São Paulo, based on information from Population and School Censuses, Household Travel Surveys and electronic transit card records.
The results illustrate how existing inequalities are amplified by variable accessibility to schools across income groups and geographical space. 
We conclude that extending the concept of local accessibility indicators to the 'school commute' can guide the development of policies to reduce educational inequalities.

<!-- # Introduction

@@

Travel to school is an everyday reality for millions of young people around the world.
The mode, duration, safety, comfort and pollution levels of this trip has huge
impacts on the future generation, yet has received relatively little academic
attention.

Travel options are vital for ensuring a more equitable supply of
educational opportunity to diverse groups. Conversely, poor accessibility
to deprived area can reinforce social inequalities, with long-term implications.
Based on this emerging context, this paper develops a school accessibility index
for local areas.

@@

This early work made the distinction between accessibility indexes that apply to zones or single 'desire lines':
"relative accessibility is defined as a measure of the effort of overcoming spatial separation between two points, while the integral accessibility is defined as a measure of the effort of overcoming spatial separation between a point and all other points within an area"
[@allen1993accesibility].

@@

-->
